Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-06-01 EMNLP-demo-paper/intro.tex

Deep neural networks have been successfully applied to various natural language processing tasks,
%where end-to-end training become a primary option.
and as the design of neural networks evolves, there is a clear trend of increasing model complexity in terms of both the architecture and the parameters count. 
%
The expressiveness of these models helps improve the prediction performance, there is a fundamental lack of interpretability leads many researchers to consider neural network models black boxes. 
%
Despite the difficulties, interpreting the model internals and reason about predictions are essential for understanding the limitation of the models and improving upon the existing designs.

@@

Recently, the attention networks have become a widely-adopted mechanism \cite{bahdanau2014neural,seo2016bidirectional,Parikh2016, VaswaniShazeerParmar2017}. The attention not only improves the model performance but also yields interpretable intermediate representations (i.e., alignment among words). As a result, attention is often considered as a natural interface for analyzing the internals of neural networks. Conduct analysis directly on the raw attention values can be challenging for human users, and therefore, visual representations that highlight word alignments between sentences have been proposed, such as bipartite graph and heatmap of the attention matrix \cite{LiChenHovy2015, li2016understanding, lee2017interactive}.  However, many challenges remain. Firstly, for some NLP tasks, the word sequences pair for which the attention is computed can be highly asymmetrical (i.e., in machine comprehension, the context paragraph can be much longer than the question sentence), which the standard visual encoding cannot adequately handle.  Also, many previous works present the attention in a static setting. However, the ability to provide an interactive environment, in which the user can instantaneously look at how changes in the input affect the attention, and how small variations in the attention alter the prediction, is crucial for interpreting the model. Finally, many previous visualization efforts, despite revealing many exciting results, focus more on illustrating potentially useful visualization technique than on providing a flexible software tool that can be easily integrated into an existing code base. As a result, applying these proposed techniques to real-world examples may involve substantial engineering effort that could prevent wide adoption.

@@

To address these challenges in interpreting attention-based models, 
we introduce a Python library that allows users to create web-based interactive visual analytics environments,
in which they can interrogate the model by perturbing the input text and observing the effects on the attention and prediction, or modifying the attention and studying how it affects predictions (see Figure~\ref{fig:modelPipeline}).
%
We demonstrate our system on two primary NLP tasks: natural language
inference (NLI) and machine comprehension (MC). 
%For the NLI task, we visualize
%decomposable attention networks which represents a series of strongly performing models \cite{Parikh2016}.
%For the MC task, we present your visualization methods on bidirectional attention flow model,
%which also represents a line of state-of-the-art models \cite{Seo2016}.

>>>>

Deep neural networks have been successfully applied to various natural language processing tasks.
%where end-to-end training become a primary option.
As the design of neural networks evolves, there is a clear trend of increasing model complexity in terms of both the architecture and the parameter count. 
%
While expressive models help improve the prediction performance, the fundamental lack of interpretability leads many researchers to consider neural network models black boxes. 
%
The ability to interpret model internals and reason about predictions is essential for understanding the limitation of the models and improving upon them.

@@

Recently, the attention networks have become  widely-adopted~\cite{bahdanau2014neural,seo2016bidirectional,Parikh2016, VaswaniShazeerParmar2017}. It not only improves the model performance but also yields interpretable intermediate representations (i.e., alignment among words). As a result, attention can provide a natural interface for analyzing the internals of neural networks. Conducting analysis directly on the raw attention values can be challenging for human users, and therefore, visual representations that highlight word alignments between sentences have been proposed, such as bipartite graph and heatmap of the attention matrix \cite{LiChenHovy2015, li2016understanding, lee2017interactive}.  However, many challenges remain. Firstly, for some NLP tasks, the word sequences pair for which the attention is computed can be highly asymmetrical (i.e., in machine comprehension, the context paragraph can be much longer than the question sentence), which the standard visual encoding cannot adequately handle.  Also, many previous works present the attention in a static setting. However, the ability to provide an interactive environment, in which the user can instantaneously look at how changes in the input affect the attention, and how small variations in the attention alter the prediction, is crucial for interpreting the model. Finally, many previous visualization efforts, despite revealing many exciting results, focus more on illustrating potentially useful visualization techniques than on providing a flexible software tool that can be easily integrated into an existing code base. As a result, applying these proposed techniques to real-world examples may involve substantial engineering effort that could prevent wide adoption.

@@

To address these challenges in interpreting attention-based models, 
we introduce a Python library that allows users to create web-based interactive visual analytics environments,
in which they can interrogate the model by perturbing the input text and observing the effects on the attention and prediction, or modifying the attention and studying how it affects predictions (see Figure~\ref{fig:modelPipeline}).
%
We demonstrate our library applied to two NLP tasks: natural language
inference (NLI) and machine comprehension (MC). 
%For the NLI task, we visualize
%decomposable attention networks which represents a series of strongly performing models \cite{Parikh2016}.
%For the MC task, we present your visualization methods on bidirectional attention flow model,
%which also represents a line of state-of-the-art models \cite{Seo2016}.

### 2018-06-01 EMNLP-demo-paper/system.tex

\section{Visualization System}
As illustrated in Figure~\ref{fig:modelPipeline}, many recent end-to-end NLP models follow a similar encoder--attention--classifier structure. The attention stage provides a window to peek into the model decision making process. 
However, the static attention alone does not tell the whole story. 
Our proposed system uses a perturbation-driven exploration strategy that allows the user manually or automatically perturb part of the pipeline and observe the changes downstream.
%
The system consists of several visualization components that can be selectively combined to form a functional interactive system for a given task. In the following section, we will cover some of the essential components; the usage of all components can be found in the accompanying video \href{https://www.youtube.com/watch?v=dcpTyWfBhp0}{link}.
%such as a prediction summary statistics the entire development set, as well as a scheme to update the model based

>>>>

\section{Visualization System}
As illustrated in Figure~\ref{fig:modelPipeline}, many recent end-to-end NLP models follow a similar encoder--attention--classifier structure. The attention stage provides a window to peek into the model decision making process. 
However, the static attention alone does not tell the whole story. 
Our proposed system uses a perturbation-driven exploration strategy that allows the user manually or automatically perturb part of the pipeline and observe the changes downstream.
%
The system consists of several visualization components that can be selectively combined to form a functional interactive system for a given task. In the following section, we will cover some of the essential components; the usage of all components can be found in the accompanying video \href{https://www.youtube.com/watch?v=PKiM4i0oIuY}{link}.
%such as a prediction summary statistics the entire development set, as well as a scheme to update the model based
